---
title: "Data Wrangling"
author: "Arash Amiririgi"
---

::: callout-note
A reduced data set related to patent data provided by the USPTO is analysed here.
:::

``` {r}
library(tidyverse)
library(data.table)
library(vroom)
library(lubridate)

col_patent_types <- list(id = col_character(), date = col_date("%Y-%m-%d"),
                         num_claims = col_double())
col_patent_assignee_types <- list(patent_id = col_character(),
                                  assignee_id = col_character())
col_assignee_types <- list(id = col_character(), type = col_double(),
                           organization = col_character())
col_uspc_types <- list(patent_id = col_character(), mainclass_id = col_double(),
                       sequence = col_double())
patent_tbl <- vroom(file = "patent.tsv", delim = "\t", col_types  = col_patent_types, 
                    na = c("", "NA", "NULL"))
patent_assignee_tbl <- vroom(file = "patent_assignee.tsv", delim = "\t",
                     col_types  = col_patent_assignee_types, na = c("", "NA", "NULL"))
assignee_tbl <- vroom(file = "assignee.tsv", delim = "\t",
                      col_types  = col_assignee_types, na = c("", "NA", "NULL"))
uspc_tbl <- vroom(file = "uspc.tsv", delim = "\t", col_types  = col_uspc_types,
                  na = c("", "NA", "NULL"))
```

## Question:
### Patent Dominance: What US company / corporation has the most patents? List the 10 US companies with the most assigned/granted patents.

### using dplyr
```{r}
patent_organization_dplyr <- left_join(patent_assignee_tbl,assignee_tbl,
                                       by= c("assignee_id"="id"))
patent_organization_dplyr %>% filter(!(is.na(organization)) & type == 2) %>%
  group_by(organization) %>%
summarise(number_of_patents = n()) %>% ungroup() %>% arrange(desc(number_of_patents)) %>%
  head(10)
```

### using data.table
```{r}
setDT(patent_assignee_tbl)
setDT(assignee_tbl)
patent_organization_dt <- patent_assignee_tbl[assignee_tbl, on= c("assignee_id"="id")]
patent_organization_dt[!(is.na(organization)) & type == 2, .N, by = organization][
  order(-N)] %>% head(10)
```

## Question:
### Recent patent activity: What US company had the most patents granted in August 2014? List the top 10 companies with the most new granted patents for August 2014.

### using dplyr
```{r}
patent_organization_date_dplyr <- left_join(patent_organization_dplyr, patent_tbl,
                                            by = c('patent_id' = 'id')) %>% select(1:5)
patent_organization_date_dplyr %>% filter(!(is.na(organization)) &
                                            month(date) == 8 & type == 2) %>%
  group_by(organization) %>% summarise(number_of_patents = n()) %>%
  ungroup() %>% arrange(desc(number_of_patents)) %>% head(10)
```

### using data.table
```{r}
setDT(patent_tbl)
patent_organization_date_dt <- patent_tbl[patent_organization_dt,
                                          on = c('id' = 'patent_id')][,!('num_claims')]
patent_organization_date_dt[!(is.na(organization)) & month(date) == 8 &
                              type == 2, .N, by = organization][order(-N)] %>% head(10)
```

## Question:
### Innovation in Tech: What is the most innovative tech sector? For the top 10 companies (worldwide) with the most patents, what are the top 5 USPTO tech main classes?

### using dplyr
```{r}
patent_organization_class_dplyr <- left_join(patent_organization_dplyr,uspc_tbl,
                               by = 'patent_id', relationship = "many-to-many") %>% select(1:5)
patent_organization_class_dplyr <- patent_organization_class_dplyr %>%
  filter(!(is.na(organization))) %>% group_by(organization) %>%
  mutate(number_of_patents = n()) %>% ungroup()
top_organization_dplyr <- patent_organization_class_dplyr %>%
  group_by(organization) %>% slice(1) %>% ungroup() %>%
  arrange(desc(number_of_patents)) %>% head(10) %>% pull(organization)
patent_organization_class_dplyr %>% filter(organization %in% top_organization_dplyr &
  !(is.na(mainclass_id))) %>% group_by(mainclass_id) %>%
  summarise(number_of_incidents = n()) %>% ungroup() %>%
  arrange(desc(number_of_incidents)) %>% head(5)
```

### using data.table
```{r}
setDT(uspc_tbl)
patent_organization_class_dt <- uspc_tbl[patent_organization_dt,
                                         on = 'patent_id'][,!('sequence')]
patent_organization_class_dt <- patent_organization_class_dt[!(is.na(organization)),
                                         number_of_patents := .N, by = organization]
top_organization_dt <- patent_organization_class_dt[, .SD[1], by = organization][
  order(-number_of_patents)] %>% head(10) %>% pull(organization)
patent_organization_class_dt[!(is.na(mainclass_id)) & organization %in% top_organization_dt,
                             .N, by = mainclass_id][order(-N)] %>% head(5)
```




